---
title: "1. Cohort selection"
output: html_notebook
---

This notebook creates the necessary materialized views in SQL and the necessary CSV files for further analyses.

```{r}
library(tidyverse)
library(RPostgreSQL)

driver <- dbDriver("PostgreSQL")
db <- dbConnect(
  driver,
  dbname = "mimiciii",
  host = "localhost",
  port = "5432",
  user = "mimicuser"
)

getSQL <- function(filepath){
  con = file(filepath, "r")
  sql.string <- ""

  while (TRUE){
    line <- readLines(con, n = 1)

    if ( length(line) == 0 ){
      break
    }

    line <- gsub("\\t", " ", line)

    if(grepl("--",line) == TRUE){
      line <- paste(sub("--","/*",line),"*/")
    }

    sql.string <- paste(sql.string, line)
  }

  close(con)
  return(sql.string)
}
```

# Generate materialized views

## Echo data

```{r}
query <- getSQL("mimic-code/concepts/echo-data.sql")
print("Generating table 'echodata'...")
dbGetQuery(db, query)
print("... Done")
```

## First day data

```{r}
query <- getSQL("mimic-code/concepts/firstday/urine-output-first-day.sql")
print("Generating table 'uofirstday'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/firstday/vitals-first-day.sql")
print("Generating table 'vitalsfirstday'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/firstday/gcs-first-day.sql")
print("Generating table 'gcsfirstday'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/firstday/labs-first-day.sql")
print("Generating table 'labsfirstday'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/firstday/blood-gas-first-day.sql")
print("Generating table 'bloodgasfirstday'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/firstday/blood-gas-first-day-arterial.sql")
print("Generating table 'bloodgasfirstdayarterial'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/firstday/ventilation-first-day.sql")
print("Generating table 'ventfirstday'...")
dbGetQuery(db, query)
print("... Done")
```


```{r}
query <- getSQL("mimic-code/concepts/firstday/height-first-day.sql")
print("Generating table 'heightfirstday'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/firstday/weight-first-day.sql")
print("Generating table 'weightfirstday'...")
dbGetQuery(db, query)
print("... Done")
```

## Demographics

```{r}
query <- getSQL("mimic-code/concepts/demographics/icustay-detail.sql")
print("Generating table 'icustay_detail'...")
dbGetQuery(db, query)
print("... Done")
```

## Durations

```{r}
query <- getSQL("mimic-code/concepts/durations/weight-durations.sql")
print("Generating table 'weightdurations'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/durations/ventilation-durations.sql")
print("Generating table 'ventdurations'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/durations/vasopressor-durations.sql")
print("Generating table 'vasopressordurations'...")
dbGetQuery(db, query)
print("... Done")
```

## Severity scores

```{r}
query <- getSQL("mimic-code/concepts/severityscores/sofa.sql")
print("Generating table 'sofa'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/severityscores/sapsii.sql")
print("Generating table 'sapsii'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/severityscores/apsiii.sql")
print("Generating table 'apsiii'...")
dbGetQuery(db, query)
print("... Done")
```

## Diagnosis categories

```{r}
query <- getSQL("mimic-code/concepts/diagnosis/ccs_diagnosis_table.sql")
print("Generating tables 'ccs_single_level_dx' and 'ccs_multi_level_dx'...")
dbGetQuery(db, query)
print("... Done")
```

## Renal replacement therapy

```{r}
query <- getSQL("mimic-code/concepts/rrt.sql")
print("Generating table 'rrt'...")
dbGetQuery(db, query)
print("... Done")
```

## Comorbidities

```{r}
query <- getSQL("mimic-code/concepts/comorbidity/elixhauser-ahrq-v37-with-drg.sql")
print("Generating table 'elixhauser_ahrq'...")
dbGetQuery(db, query)
print("... Done")
```

## Norepinephrine

```{r}
query <- getSQL("mimic-code/concepts/durations/norepinephrine-dose.sql")
print("Generating table 'norepinephrine_dose'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("mimic-code/concepts/duratins/norepinephrine-durations")
print("Generating table 'norepinephrinedurations'...")
dbGetQuery(db, query)
print("... Done")
```

## Suspicion of infection

```{r}
query <- getSQL("sql/suspected_infection.sql")
print("Generating table 'suspected_infection'...")
dbGetQuery(db, query)
print("... Done")
```

## Arterial blood gases

```{r}
query <- getSQL("sql/all_abg.sql")
print("Generating table 'abg_all'...")
dbGetQuery(db, query)
print("... Done")
```

## Infectious sources

```{r}
query <- getSQL("sql/infectious_sources.sql")
print("Generating table 'infectious_sources'...")
dbGetQuery(db, query)
print("... Done")
```

## Cohort tables

```{r}
query <- getSQL("sql/cohort.sql")
print("Generating table 'cohort_all'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("sql/abg_sepsis.sql")
print("Generating table 'abg_sepsis'...")
dbGetQuery(db, query)
print("... Done")
```

```{r}
query <- getSQL("sql/cohort_sepsis.sql")
print("Generating table 'cohort_sepsis'...")
dbGetQuery(db, query)
print("... Done")
```

# Cohort selection

```{r}
cohort_all <- tbl(db, "cohort_all")
manual_review <- readLines("data/not_sepsis.csv")
```

## All ICU admissions

```{r}
cohort_all %>%
  tally()
```

## Adult patients

```{r}
cohort_all %>%
  filter(admission_age >= 18) %>%
  tally()
```

## First ICU stay

```{r}
cohort_all %>%
  filter(admission_age >= 18 & 
           icustay_seq == 1) %>%
  tally()
```

## LOS at least 24 hours

```{r}
cohort_all %>%
  filter(admission_age >= 18 & 
           icustay_seq == 1 &
           los_icu >= 1) %>%
  tally()
```

## Exclude CSRU, CCU and TSICU patients

```{r}
cohort_all %>%
  filter(admission_age >= 18 & 
           icustay_seq == 1 &
           los_icu >= 1 &
           !first_careunit %in% c("CSRU", "CCU", "TSICU")) %>%
  tally()
```

## At least 1 ABG

```{r}
cohort_all %>%
  filter(admission_age >= 18 & 
           icustay_seq == 1 &
           los_icu >= 1 &
           !first_careunit %in% c("CSRU", "CCU", "TSICU") &
           abg_count >= 1) %>%
  tally()
```

## At least 24h of mechanical ventilation during the first 24 hours

```{r}
cohort_all %>%
  filter(admission_age >= 18 & 
           icustay_seq == 1 &
           los_icu >= 1 &
           !first_careunit %in% c("CSRU", "CCU", "TSICU") &
           abg_count >= 1 &
           vent_first_duration >= 24) %>%
  tally()
```

## Sepsis patients

```{r}
cohort_all %>%
  filter(admission_age >= 18 & 
           icustay_seq == 1 &
           los_icu >= 1 &
           !first_careunit %in% c("CSRU", "CCU", "TSICU") &
           abg_count >= 1 &
           vent_first_duration >= 24 &
           flag_sepsis == 1) %>%
  tally()
```

## Septic shock

```{r}
cohort_all %>%
  filter(admission_age >= 18 & 
           icustay_seq == 1 &
           los_icu >= 1 &
           !first_careunit %in% c("CSRU", "CCU", "TSICU") &
           abg_count >= 1 &
           vent_first_duration >= 24 &
           flag_sepsis == 1 &
           flag_shock == 1) %>%
  tally()
```

## Final cohort

```{r}
cohort_all %>%
  filter(admission_age >= 18 & 
           icustay_seq == 1 &
           los_icu >= 1 &
           !first_careunit %in% c("CSRU", "CCU", "TSICU") &
           abg_count >= 1 &
           vent_first_duration >= 24 &
           flag_sepsis == 1 &
           flag_shock == 1 &
           !subject_id %in% !!manual_review) %>%
  tally()
```

# Datasets

## Sepsis cohort

```{r}
cohort_sepsis <- tbl(db, "cohort_sepsis") %>% collect()

write_csv(cohort_sepsis, "data/cohort_sepsis.csv")
```

## Septic shock cohort 
```{r}
cohort_shock <- cohort_sepsis %>%
  filter(flag_shock == 1 &
           !subject_id %in% !!manual_review)

write.csv(cohort_shock, "data/cohort_shock.csv")
```

```{r}
dbDisconnect(db)
dbUnloadDriver(driver)
```
